---
title: "Survival Analysis NUP98-KDM5A and CBFGLIS"
author: "Jenny Smith"
date: "July 19, 2017"
output: html_document
---


```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center')
knitr::opts_knit$set(root.dir = '~/RNA_seq_Analysis/2017.06.20_DEGs_RareVariants/')
```


```{r message = FALSE, warning=FALSE}
library(stringr)
library(magrittr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(plyr)
getwd()
```

```{r}
source("~/scripts/RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/DifferentialExpressionPipeline.r")
source("~/scripts/survival_analysis/Survplot_Functions_2017.03.31.r")
source("~/scripts/RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/Heatmaps_Function.r")
```


#Read in the clinical Data

```{r}
RAM <- read.csv("~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/New_stratification_outcomes_updated_2017.03.28.csv", stringsAsFactors = FALSE, row.names = 1)
head(RAM[,c(1,4,8)])
```


```{r}
clinData <- read.csv("~/reference_mapping-files/TARGET_AML_current_asof_june30_2016_UPDATED_RhondaEdit_2017.06.08.csv", stringsAsFactors = FALSE, row.names=1)

dim(clinData)
```


```{r}
additionalGLIS <- read.csv("~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/CBFGLIS_from0531_from03P1.csv", stringsAsFactors = FALSE)
additionalGLIS <- additionalGLIS[21:27,]

# dim(additionalGLIS)
head(additionalGLIS)
# additionalGLIS
```


```{r}
clinData <- rbind.fill(clinData, additionalGLIS)

dim(clinData)
```


```{r}
head(clinData[,c(1, 10:12)])
```


```{r}
# write.csv(clinData, "~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/TARGET_AML_current_asof_june30_2016_withCBF-GLIS_2017.06.08.csv")
```

```{r}
table(clinData$Fusion.Cluster.Category) #27 CBF-GLIS
```

#Select Columns 

```{r}
colnames(clinData)
```

```{r}
cols <- colnames(clinData)[28:45]
# cols
```

```{r}
#Check that they do not have different factor levels besides No, Yes, Unknown
table(sapply(clinData[,cols], unique) )
```







#Select Fusions 

Note: DEK-NUP214 is t(6;9)
Inv(16) is CBFB-MYH11
t(8;21) is RUNX1-RUNX1T1

```{r}
# clinData$Known.Fusions <- gsub("\\s", "", clinData$Known.Fusions) #remove exra whitespace
unique(clinData$Reason)
# unique(clinData$Fusion.Cluster.Category)
```

```{r}
#Select fusions of interest
fusions <- c("NUP98-KDM5A", "NUP98-NSD1","NUP98-HOXD13","NUP98-PHF23","NUP98-HMGB3", "NPM1","MECOM","KAT6A Fusion","ETS Fusion")


#Define the phenotype vectors
fusionAnnotations <- as.data.frame(lapply(fusions, function(fusion) 
  ifelse(grepl(fusion, clinData$Reason), gsub("\\-|\\s", "\\.",fusion), "otherAML") %>% 
                                            set_names(clinData$TARGET.USI.1)), stringsAsFactors=FALSE)

colnames(fusionAnnotations) <- gsub("\\-|\\s", "\\.", fusions)

# head(fusionAnnotations)
dim(fusionAnnotations)
sapply(fusionAnnotations, unique)
```


```{r}
unique(clinData$Fusion.Cluster.Category)
```

```{r}
#Select fusions of interest
fusions2 <- c("CBFA2T3-GLIS2","NUP98-X")

#Define the phenotype vectors
fusionAnnotations2 <- as.data.frame(lapply(fusions2, function(fusion) ifelse(grepl(fusion, clinData$Fusion.Cluster.Category), gsub("\\-|\\s", "\\.",fusion), "otherAML") %>%
                                             set_names(clinData$TARGET.USI.1)), stringsAsFactors=FALSE)

colnames(fusionAnnotations2) <- gsub("\\-|\\s", "\\.", fusions2)

# head(fusionAnnotations2)
dim(fusionAnnotations2)
sapply(fusionAnnotations2, unique)
```

```{r}
fusionAnnotations <- transform(merge(fusionAnnotations, fusionAnnotations2, by.x=0, by.y=0), row.names = Row.names, Row.names=NULL) 
dim(fusionAnnotations)
# head(fusionAnnotations)
```

```{r}
RAM.vector  <- ifelse(grepl("RAM", RAM$Reason), "RAM", "otherAML") %>% set_names(., rownames(RAM))
# RAM.vector
```


```{r}
#merge clincal data with the fusion data

clinData.withFusions <- merge(clinData, fusionAnnotations, 
                              by.x="TARGET.USI.1", by.y=0, stringsAsFactors=FALSE) %>% 
  set_rownames(., .[,"TARGET.USI.1"])

clinData.withFusions <- merge(clinData.withFusions, RAM.vector, by.x=0, by.y=0, all.x = TRUE, stringsAsFacors=FALSE) %>% 
  set_rownames(., .[,"TARGET.USI.1"])

colnames(clinData.withFusions)[which(colnames(clinData.withFusions) == "y")] <- "RAM"

clinData.withFusions$RAM <- ifelse(is.na(clinData.withFusions$RAM), "otherAML", as.character(clinData.withFusions$RAM))
```


```{r}
#Cleanup the data
clinData.withFusions$Fusion.Cluster.Category <- ifelse(clinData.withFusions$Fusion.Cluster.Category == "", clinData.withFusions$Known.Fusions, clinData.withFusions$Fusion.Cluster.Category)

clinData.withFusions$Bone.marrow.leukemic.blast.percentage.... <- as.numeric(as.character(ifelse(clinData.withFusions$Bone.marrow.leukemic.blast.percentage.... == ".", NA, clinData.withFusions$Bone.marrow.leukemic.blast.percentage....)))

clinData.withFusions$OS.ID <- ifelse(clinData.withFusions$Vital.Status == "Alive", 0, 1)
clinData.withFusions$Event.ID <- ifelse(clinData.withFusions$First.Event == "Censored", 0, 1)


# dim(clinData.withFusions) # 993  by 91
head(clinData.withFusions[,c(76,77,78,79)])
# write.csv(clinData.withFusions, "TARGET_AML_CDEs_withFusionCols_11.15.2017.csv")
```


```{r}
table(clinData.withFusions$CBFA2T3.GLIS2)
```



#Survival Analysis for AAML0531

#Overall Survival 
```{r message=FALSE}
clinData.withFusions$NUP98.KDM5A <- relevel(as.factor(clinData.withFusions$NUP98.KDM5A), ref="otherAML")
clinData.withFusions$CBFA2T3.GLIS2 <- relevel(as.factor(clinData.withFusions$CBFA2T3.GLIS2), ref="otherAML")

Fus.Names <- c("NUP98.KDM5A", "CBFA2T3.GLIS2")
OS <- lapply(c("NUP98.KDM5A", "CBFA2T3.GLIS2"), FUN = SurvObjects,df=clinData.withFusions,
             colNames=c("Overall.Survival.Time.in.Days","OS.ID"),
             time="DtoY") 
names(OS) <- Fus.Names
# lapply(OS, function(x) x[2:4])
```


```{r}
summary(OS$CBFA2T3.GLIS2$CoxPH)
# OS$CBFA2T3.GLIS2$survFit
```

```{r}
# summary(OS$NUP98.KDM5A$CoxPH)
```

```{r}
# summary(OS$NUP98.KDM5A$survFit, time=c(0,3,5))
```


#Kaplan-Meier Plots

```{r}
OS.fit <- list(OS$NUP98.KDM5A$survFit, OS$CBFA2T3.GLIS2$survFit)
NUPColors <- c(NUP98.KDM5A="firebrick1", otherAML="lightskyblue4")
CBFColors <- c(CBFA2T3.GLIS2="red", otherAML="darkblue")
colors <- list(NUPColors, CBFColors)
```


```{r message=FALSE}
OS.plots <- mapply(FUN=SurvivalPlot, OS.fit, colors, 
                   MoreArgs = list(LegendTitle="Overall Survival 0531", timeUnit="Years"), SIMPLIFY = FALSE)

names(OS.plots) <- Fus.Names
```

```{r message=FALSE}
# pdf(file="TARGET_AML_NUP98KDM5A_vs_OtherAML_OSplot.pdf", height = 5, width = 7)
OS.plots$NUP98.KDM5A + xlim(0,6)
# dev.off()

OS.plots$CBFA2T3.GLIS2
```

```{r}
# lapply(names(OS.plots), function(i) ggsave(filename=paste0("OS_",i,".pdf"), plot=OS.plots[[i]], device = "pdf", width = 7, height = 5, unit="in"))
```


#Event-Free Survival 

```{r}
EFS <- lapply(c("NUP98.KDM5A", "CBFA2T3.GLIS2"), FUN = SurvObjects,df=clinData.withFusions,
             colNames=c("Event.Free.Survival.Time.in.Days","Event.ID"),
             time="DtoY") 
names(EFS) <- Fus.Names

# lapply(EFS, function(x) x[2:4])
```

```{r}
# EFS$CBFA2T3.GLIS2$survFit
# EFS$CBFA2T3.GLIS2$log.rank
```


#Kaplan-Meier Plots

```{r message=FALSE}
EFS.fit <- list(EFS$NUP98.KDM5A$survFit, EFS$CBFA2T3.GLIS2$survFit)

EFS.plots <- mapply(FUN=SurvivalPlot, EFS.fit, colors, 
                   MoreArgs = list(LegendTitle="Event-Free Survival 0531", timeUnit="Years"), 
                   SIMPLIFY = FALSE) 
names(EFS.plots) <- Fus.Names
EFS.plots
```

```{r}
# lapply(names(EFS.plots), function(i) ggsave(filename=paste0("EFS_",i,".pdf"), plot=EFS.plots[[i]], device = "pdf", width = 7, height = 5, unit="in"))
```


#Survival Analysis on 1031

```{r}
CDE.1031 <- read.csv("~/reference_mapping-files/TARGET_AML_AAML1031_merged_CDE_Cleaned_28Sept2017.csv",
                     stringsAsFactors = FALSE,row.names = 1,
                     na.strings = c("","#N/A"," ","NA"))

dim(CDE.1031)
```


```{r}
head(CDE.1031)
```


```{r message=FALSE, warning=FALSE}
#this code was used to create TARGET_AML_AAML1031_merged_CDE_Cleaned_20Sept2017.csv. Not needed now
# CDE.1031[,6:55] <- sapply(CDE.1031[,6:55], function(x) ifelse(is.na(x), "Not available", x))
# CDE.1031[,6:55] <- sapply(CDE.1031[,6:55], function(x) ifelse(x=="Not evaluated", "Not available", x))
# CDE.1031[,c(7,10,56:63)] <- sapply(CDE.1031[,c(7,10,56:63)], function(x) as.numeric(as.character(x)))
# dim(CDE.1031)
```

```{r}
# write.csv(CDE.1031, "~/reference_mapping-files/TARGET_AML_AAML1031_merged_CDE_Cleaned_20Sept2017.csv")
# save(CDE.1031, file="TARGET_AML_1031_CDE_cleaned_20Sept2017.csv")
```


```{r}
CBFGLIS.1031 <- read.csv("~/reference_mapping-files/TARGET_AML_CBFGLIS_1031dx5&underfusiondataFINAL.csv", stringsAsFactors = FALSE)
# dim(CBFGLIS.1031) #only 288 were tested
```


```{r}
# unique(CBFGLIS.1031$CBFT2A3.GLIS.Y.N) 
pos.1031 <- CBFGLIS.1031[CBFGLIS.1031$CBFT2A3.GLIS.Y.N == "YES",]$REG.
neg.1031 <- CBFGLIS.1031[CBFGLIS.1031$CBFT2A3.GLIS.Y.N == "",]$REG.

length(pos.1031) #13
length(neg.1031)
```

```{r}
RAM <- read.csv("~/reference_mapping-files/TARGET_AML_RAM_1031_05112017.csv", stringsAsFactors = FALSE)
head(RAM[,1:3])
```

```{r message=FALSE, warning=FALSE}
RAM.Clean <- RAM %>%
  separate(Patient.registration.number, sep="[ \\-]", into=c("Reg", "Initials"))

head(RAM.Clean[,1:3])
```


```{r}
CDE.1031 <- CDE.1031 %>% 
  mutate(CBFGLIS.Status=ifelse(Patient.ID %in% pos.1031, "Yes", ifelse(Patient.ID %in% neg.1031, "No", "Unknown"))) %>% #13 CBF-GLIS of 13 are in CDE
  mutate(RAM.Status=ifelse(Patient.ID %in% RAM.Clean$Reg, "Yes", "No")) %>%  #16 RAM of the 33 have CDE information
  mutate(Age.yrs=Age..Days./365.24)

dim(CDE.1031) #1559 
head(CDE.1031[,c("CBFGLIS.Status", "RAM.Status", "Age.yrs")])
```

```{r}
# write.csv(CDE.1031, "TARGET_AML_1031_CDE_cleaned_11.15.2017.csv", row.names = FALSE)
```



```{r}
#Clean the clinical data
CDE.1031.withOS <- CDE.1031 %>% 
  mutate(CBFGLIS.Status=ifelse(Patient.ID %in% pos.1031, "yes", "no")) %>% #13 CBF-GLIS of 13 are in CDE
  mutate(RAM.Status=ifelse(Patient.ID %in% RAM.Clean$Reg, "yes", "no")) %>%  #16 RAM of the 33 have CDE information
  filter(! is.na(osi)) %>%
  mutate(Age.yrs=Age..Days./365.24)
  
head(CDE.1031.withOS)
```


```{r}
# write.csv(CDE.1031, "~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/TARGET_AML_1031_CDE_cleaned_20Sept2017.csv", row.names = FALSE)
```


```{r}
dim(CDE.1031.withOS)#667 patients 
```

```{r}
os.cols <- c("yrsos", "osi")
efs.cols <- c("yrsefs", "efsi")
```

```{r}
OS.1031 <- SurvObjects(CDE.1031.withOS, os.cols, "CBFGLIS.Status")
OS.1031$survFit
OS.1031$Gehan.Breslow.Wilcoxon
```

```{r}
EFS.1031 <- SurvObjects(CDE.1031.withOS, efs.cols, "CBFGLIS.Status")
EFS.1031$survFit
EFS.1031$log.rank
```


#Plot the Kaplan-Meiers

```{r}
os.plot <- SurvivalPlot(OS.1031$survFit, "CBFA2T2-GLIS2", "Years", c(yes="wheat4", no="wheat3"))
os.plot <- os.plot + labs(title="AAML1031: Overall Survival") + geom_text(x=4.5,y=0.2, label="p < 0.001", size=5)
os.plot

# ggsave(filename = "OS_1031_CBFGLISvsOtherAML_KMplot.pdf", plot = os.plot,
       # device = "pdf", height = 5, width = 7, units = "in")
```

```{r}
efs.plot <- SurvivalPlot(EFS.1031$survFit, "CBFA2T3-GLIS2", "Years", c(yes="wheat4", no="wheat3"))
efs.plot <- efs.plot + labs(title="AAML1031:Event-Free Survival") + geom_text(x=4.5, y=0.2, label="p < 0.001", size=5)
efs.plot

# ggsave(filename = "EFS_1031_CBFGLISvsOtherAML_KMplot.pdf", plot = efs.plot, 
#        device = "pdf", height = 5, width = 7, units = "in")
```



#Clinical Characterisitcs 


## AAML0531 Clinical Characteristics 

```{r message=FALSE}
library(compareGroups)
```

```{r}
clinCols <- c(
  "Age.at.Diagnosis.in.Days",
  "Gender", 
  "WBC.at.Diagnosis",
  "Bone.marrow.leukemic.blast.percentage....",
  "Primary.Cytogenetic.Code",
  "Fusion.Cluster.Category",
  "t.8.21.",
  "inv.16.",
  "FLT3.ITD.positive.", 
  "WT1.mutation",
  "NPM.mutation",
  "RAM", 
  "NUP98.NSD1", 
  "CBFA2T3.GLIS2", 
  "KAT6A.Fusion", 
  "NUP98.KDM5A"
)
```

```{r}
ageCats <- function(Age.Vector){
  #age vector is a named vector with the patient ages. 
  
  ageCat <- function(age){
    if (age < 5){
      g <-"under5"
    }else if (age >= 5 & age < 11 ){
      g <- "Age5to10"
    }else if (age >= 11 & age < 16){
      g <- "Age11to15"
    }else if (age >=16){
      g <- "AgeGreater15"
    }
  }

  ageCategories <- sapply(Age.Vector, ageCat)
  
  return(ageCategories)
}
```

```{r}
sub <- clinData.withFusions[,clinCols]
sub$Age.Years <- sub$Age.at.Diagnosis.in.Days/365
sub <- sub[,-1] 

ages <- sub$Age.Years
sub$Age.Categories <-  ageCats(ages)

head(sub[,1:5])
```


#CBFGLIS2 in AAML0531
```{r}
comp.CBF <- compareGroups(CBFA2T3.GLIS2 ~ . ,data = sub, method=4, Q1=0, Q3=1)

tab.CBF <- createTable(comp.CBF)

tab.CBF
export2csv(tab.CBF, "TARGET_AML_CBFGLIS_AAML0531_CDE_Table.csv")
```


#CBFGLIS with Flow Data

```{r}
flow <- read.csv("~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/FlowCytometry/TARGET_AML_CBFGLIS_PatientUSI_27June2017-2017-7-18FD_FLOWcytometetryData.csv", stringsAsFactors = FALSE, row.names = 1)
flow <- flow[complete.cases(flow),] #removes NAs
head(flow[,1:5])
```

```{r}
dim(flow) #437 
```

```{r}
#select only CBFGLIS with FLOW data
sub.flow <- transform(merge(flow, sub, by.x=0, by.y=0), row.names=Row.names, Row.names=NULL) 
head(sub.flow[,1:5])#10 CBF-GLIS patients, 437 patients  
```

#CBF and RAM in the flow Data

```{r}
comp.CBF.flow <- compareGroups(CBFA2T3.GLIS2 ~ RAM + CD56,method=4, data = sub.flow, Q1=0, Q3=1)

createTable(comp.CBF.flow)
```


#Clinical Chars of  NUP98-KDM5A

```{r}
comp.NUP <- compareGroups(NUP98.KDM5A ~ ., method=4, Q1=0, Q3=1, data=sub)

tab.NUP <- createTable(comp.NUP)

tab.NUP
# export2csv(tab.NUP, file="TARGET_AML_NUP98.KDM5A_ClinCharacteristicsTable.csv")
```


##AAML1031 Clinical Characteristics

```{r}
# colnames(CDE.1031)
```

```{r}
forCDETable <- c("Age.yrs",
                 "Gender", 
                 "Race", 
                 "RAM.Status",
                 "CBFGLIS.Status", 
                 "FLT3.ITD", 
                 "CEBPA", 
                 "NPM1",
                 "Monosomy.5.del.5q..",
                 "Abnormal.Karyotype.Inv.16.t.16.16.",
                 "Abnormal.Karyotype..t.8.21..",
                 "Abnormal.Karyotype.Monosomy.7.",
                 "MRD.status")

dat <- CDE.1031[,forCDETable]
dim(dat)
```

```{r}
comp <- compareGroups(CBFGLIS.Status ~ ., data = dat, method=4, Q1=0, Q3=1)

tab <- createTable(comp)
tab
```

```{r}
# export2csv(tab, file = "TARGET_AAML1031_CBFGLIS_CDE_Table.csv")
# export2latex(tab, file="TARGET_AAML1031_CBFGLIS_CDE_Table.tex")
```



```{r}
CDE.1031 %>%
  filter(Cytogenetic.test.result. == "Normal" & CBFGLIS.Status == "yes") %>%
  select(USI)
```

```{r}
library(dplyr)
clinData.withFusions %>%
  filter(Primary.Cytogenetic.Code == "Normal" & CBFA2T3.GLIS2 == "CBFA2T3.GLIS2") %>%
  select(TARGET.USI.1)
```

```{r}
perCent.CytoNormal <- round((4 + 11)/33 * 100, digits=2)
perCent.CytoNormal
```

```{r}
sum(clinData.withFusions$Primary.Cytogenetic.Code == "Normal" & clinData.withFusions$CBFA2T3.GLIS2 == "otherAML") + sum(CDE.1031$Cytogenetic.test.result. == "Normal" & CDE.1031$CBFGLIS.Status == "no")
```

```{r}
345/1625
```

```{r}
table <- cbind(c(15,18), c(345, 1280)) %>% 
  set_colnames(c("Pos", "Neg")) %>%
  set_rownames(c("CN","Not CN"))

chisq.test(table)
```

```{r}
CDE.1031 %>% 
  filter(CBFGLIS.Status == "yes" & Abnormal.Karyotype.Inv.16.t.16.16. == "Positive") %>% 
  select(Karyotype)
```



#Combine 1031 and 0531 

```{r}
# colnames(clinData.withFusions)
```

```{r}
# colnames(CDE.1031)
```

```{r}
CDE.0531 <- clinData.withFusions %>%
  mutate(Age.yrs=Age.at.Diagnosis.in.Days/365.25,
         monosomy.5.del.5.q=ifelse(clinData.withFusions$del5q == "Yes" | clinData.withFusions$monosomy.5 == "Yes", "Yes", "No"),
         EFS.yrs=Event.Free.Survival.Time.in.Days/365.25,
         OS.yrs=Overall.Survival.Time.in.Days/365.25,
         CBFA2T3.GLIS2=as.character(CBFA2T3.GLIS2),
         CR1=ifelse( grepl("CR|Not in CR", clinData.withFusions$CR.status.at.end.of.course.1),clinData.withFusions$CR.status.at.end.of.course.1, NA),
         MRD.EOI1=as.numeric(as.character(MRD...at.end.of.course.1)))
```

```{r}

CDE.0531.cols <- c("TARGET.USI.1",
  "Age.yrs",
  "Gender", 
  "Race",
  "Ethnicity",
  "WBC.at.Diagnosis",
  "MRD.EOI1",
  "Risk.group",
  "t.8.21.",
  "inv.16.",
  "FLT3.ITD.positive.", 
  "NPM.mutation",
  "CEBPA.mutation",
  "monosomy.5.del.5.q",
  "monosomy.7",
  "RAM", 
  "CBFA2T3.GLIS2",
  "CR1",
  "EFS.yrs",
  "OS.yrs",
  "Event.ID",
  "OS.ID")

CDE.1031.cols <- c("USI",
                 "Age.yrs",
                 "Gender", 
                 "Race",
                 "Ethnicity",
                 "wbc1",
                 "MRD.percentage",
                 "Cytogenetic.Risk.Group.",
                 "Abnormal.Karyotype..t.8.21..",
                 "Abnormal.Karyotype.Inv.16.t.16.16.",
                 "FLT3.ITD",
                 "NPM1",
                 "CEBPA", 
                 "Monosomy.5.del.5q..",
                 "Abnormal.Karyotype.Monosomy.7.",
                 "RAM.Status",
                 "CBFGLIS.Status",
                 "cr1",
                 "yrsefs",
                 "yrsos",
                 "efsi",
                 "osi")
```

What are the units for WBC? its not percent
```{r}
CDE.1031.m <- CDE.1031.withOS[,CDE.1031.cols] %>% set_colnames(CDE.0531.cols)

merged <- rbind(CDE.0531[,CDE.0531.cols], CDE.1031.m) %>% 
  mutate(Protocol=ifelse(TARGET.USI.1 %in% clinData.withFusions$TARGET.USI.1, "AAML0531", "AAML1031"),
         WBC=as.numeric(as.character(WBC.at.Diagnosis)))

head(merged)
```

```{r}
dim(merged)
```


```{r}
sapply(merged[,c(3:5,8:18)], unique)
```


```{r}
#Fis Risk Group
merged[merged == "High Risk"] <- "High"
merged[merged == "Low Risk"] <- "Low"
merged[merged == "Standard Risk"] <- "Standard"
merged[merged == "Unable to determine" | merged == "Not available"] <- "Unknown"
merged$Risk.group <- ifelse(merged$Risk.group == "10" | merged$Risk.group == "30", "Unknown", merged$Risk.group)

#Fix Pos and Neg
merged[merged == "Negative"| merged == "no"] <- "No"
merged[merged == "Positive" | merged == "yes" ] <- "Yes"

#Fix FLT3-ITD
merged[merged == "Wild-type"] <- "No"
merged[merged == "Internal tandem duplication"] <- "Yes"

#Fix Fusion Cols
merged[merged == "otherAML"] <- "No"
merged[merged == "RAM"] <- "Yes"
merged[merged == "CBFA2T3.GLIS2"] <- "Yes"

#Fix CR
merged$CR1 <- ifelse(merged$CR1 == "0", "Not in CR", merged$CR1 )
merged$CR1 <- ifelse(merged$CR == "1", "CR", merged$CR1)
```

```{r}
sapply(merged[,c(3:5,8:18)], unique)
```

```{r}
merged <- unique(merged)
dim(merged)
```

```{r}
write.csv(merged, "~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/TARGET_AML_0531_1031_CDE_merged_11.15.17.csv", row.names = FALSE)
# save(merged, file="~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/TARGET_AML_0531_1031_CDE_merged_14Sept2017.RData")
# merged
```


```{r}
final.order <- c("TARGET.USI.1",
  "Protocol",
  "Age.yrs",
  "Gender",
  "Race",
  "Ethnicity",
  "WBC",
  "MRD.EOI1",
  "Risk.group",
  "t.8.21.",
  "inv.16.",
  "FLT3.ITD.positive.",
  "NPM.mutation",
  "CEBPA.mutation",
  "monosomy.5.del.5.q",
  "monosomy.7",
  "RAM",
  "CBFA2T3.GLIS2",
  "CR1",
  "EFS.yrs",
  "OS.yrs",
  "Event.ID",
  "OS.ID")

merged <- merged[,final.order]
colnames(merged)
```

#Survival Analysis 

```{r}
os.cols2 <- c("OS.yrs", "OS.ID" )
efs.cols2 <- c("EFS.yrs", "Event.ID")
```

#OS
```{r}
os.merged <- SurvObjects(merged, colNames = os.cols2, group="CBFA2T3.GLIS2")
os.merged$survFit
os.merged$log.rank
```

```{r}
summary(os.merged$survFit, time=c(0,3,5))
```


```{r}
summary(os.merged$CoxPH)
```


#EFS
```{r}
efs.merged <- SurvObjects(merged, colNames = efs.cols2, group="CBFA2T3.GLIS2")
efs.merged$survFit
efs.merged$log.rank
```

Check when the p-value is calculated. I beleive its when 50% of patients have an event (the median survival time)

```{r}
summary(efs.merged$survFit, time=c(0,3,5))
```


```{r}
summary(efs.merged$CoxPH)
```



#Kaplan-Meier Plots 

```{r}
osplot <- SurvivalPlot(os.merged$survFit, "CBFA2T3-GLIS2", "Years", c(Yes="red", No="darkblue")) + labs(title="Overall Survival: CBFA2T3-GLIS2") + geom_text(x=10, y=0.15, label="p < 0.001", size=5)
osplot

# ggsave(filename = "overall_survival_CBFGLIS_0531_1031_KMplot.pdf", device = "pdf",
       # height = 5, width = 7, units="in")
```

```{r}
efsplot <- SurvivalPlot(efs.merged$survFit, "CBFA2T3-GLIS2", "Years", c(Yes="red", No="darkblue")) + 
  labs(title="Event-Free Survival: CBFA2T3-GLIS2") + geom_text(x=10,y=0.1, label="p = 0.002", size=5)
efsplot

# ggsave(filename = "event-free_survial_CBFGLIS_0531_1031_KMplot.pdf", plot=efsplot, dev="pdf",
       # height=5, width = 7)

```



#Clinical Characteristics

```{r}
library(compareGroups)

comp.merge <- compareGroups(CBFA2T3.GLIS2 ~ ., data=merged[,-1], method =4 , Q1=0, Q3=1)
tab.merge <- createTable(comp.merge)
tab.merge
```


```{r}
# export2latex(tab.merge,"~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/CBFGLIS_CDE_Table_0531_1031_14Sept2017.tex")
# export2csv(tab.merge,"~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/CBFGLIS_CDE_Table_0531_1031_14Sept2017.csv")
```



#Ages 

```{r}
AgeCategory <- merged %>% 
  filter(CBFA2T3.GLIS2 == "Yes") %>% 
  mutate(AgeCat=ifelse(Age.yrs < 5, "under5", "over5")) %>%
  mutate(percent=ifelse(AgeCat=="under5", sum(AgeCat=="under5")/33*100, sum(AgeCat=="over5")/33*100))
```

```{r}
sum(AgeCategory$AgeCat == "under5")
sum(AgeCategory$AgeCat == "over5") 
```

```{r}
N.Pat.Under5 <- sum(merged$Age.yrs < 5)
N.Pat.Under5
```



#FAB-M7/t(1;22)/AMKL


```{r}
t.1.22 <- function(ISCN){
  ifelse(grepl("t\\(1;22\\)", ISCN), "yes", "no")
}

M7.0531 <- clinData.withFusions %>% 
  filter(FAB.Category == "M7") 

# table(M7.0531$CBFA2T3.GLIS2, M7.0531$FAB.Category, M7.0531$Fusion.Cluster.Category) #8 CBFGLIS are M7 and 5 are t(1:22)
```

```{r}
M7.1031 <- CDE.1031 %>%
  filter(grepl("t\\(1;22\\)",Other.recurrent.abnormalities.))


M7.1031
# table(M7.1031$CBFGLIS.Status, M7.1031$Other.recurrent.abnormalities.)
```




#Oncoprint

```{r}
library(ComplexHeatmap)
```

http://bioconductor.org/packages/release/bioc/vignettes/ComplexHeatmap/inst/doc/s8.oncoprint.html

```{r}
# merged <- unique(merged)
rownames(merged) <- merged$TARGET.USI.1
head(merged)
```


```{r}
onco.df <- merged[,10:18]

onco.list <- list()
for (col in colnames(onco.df)){
  vector <- onco.df[,col]
  binary <- ifelse(vector == "Yes", 1,0)
  mat <- t(as.matrix(binary)) %>%
    set_colnames(rownames(onco.df)) %>% 
    set_rownames(col)
  onco.list[[col]] <- mat
}

class(onco.list$t.8.21.)
dim(onco.list$t.8.21.)
onco.list$t.8.21.[,1:15]
```


```{r}
colors <- c("turquoise3", "yellow", "blue", "firebrick1", 
                "black","seagreen2", "maroon", "orchid", "cornflowerblue", 
                "darkblue", "azure4", "chartreuse1", "darkmagenta","orange1", 
                "deeppink", "darkslategray1", "green4", "navajowhite2",
                "brown3", "darkgoldenrod3", "deepskyblue1", "lightcoral", 
                "mediumorchid", "saddlebrown")
```


```{r}
groups <- names(onco.list)

cc <- NULL
for (i in 1:length(groups)){
      c <- colors[i]
      cc <- c(cc,c)
}


names(cc) <- groups
```

```{r}
testset <- c("CBFA2T3.GLIS2", "RAM","t.8.21.")
```


```{r }
alter_fun1 = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#CCCCCC", col = NA))
    },
    CBFA2T3.GLIS2 = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "blue", col = NA))
    },
    RAM = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "red", col = NA))
    },
    t.8.21. = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h*0.33, gp = gpar(fill = "#008000", col = NA))
    }
)

```



```{r fig.width=10, fig.height=10}
# cc[testset]
# oncoPrint(onco.list[testset],
          # alter_fun = alter_fun1, col=cc[testset])
```






#Try to Use Geom Tile

```{r}
tile.df <- merged %>% 
  filter(CBFA2T3.GLIS2 == "Yes") %>% 
  select_(.dots=colnames(merged)[c(1,10:18)])

head(tile.df)
```

```{r}
tile.binary <- as.data.frame(apply(tile.df[,-1], 1, function(x) ifelse(x == "Yes", 1, 0)))
colnames(tile.binary) <- tile.df$TARGET.USI.1
tile.binary <- cbind(CytGroup=rownames(tile.binary), tile.binary)

tile.binary[1:5,1:5]
```


```{r}
library(reshape2)
tile.tidy <-  tile.binary %>% melt()
head(tile.tidy)
```


```{r}
tile.w.perc <- tile.tidy %>% 
  group_by(CytGroup) %>%
  mutate(n=n()) %>%
  mutate(postiveTotal=sum(value)) %>%
  mutate(percent=sum(value)/n*100) %>%
  arrange(desc(-percent,-value))


CytoOrder <- unique(tile.w.perc$CytGroup)

tile.w.perc
# CytoOrder
```


```{r}
patients <- as.data.frame(t(tile.binary[,-1]))

#patient order is the decreasing values for variables ordered by percentage ctypGroup
patietOrder <- rownames(patients[order(patients$CBFA2T3.GLIS2,
               patients$RAM, 
               patients$CEBPA.mutation,
               patients$FLT3.ITD.positive.,
               patients$inv.16.,
               decreasing=TRUE),])
patietOrder
```




```{r fig.width=10, fig.height=4}
tile.tidy$CytGroup <- factor(tile.tidy$CytGroup, levels = CytoOrder)
tile.tidy$variable <- factor(tile.tidy$variable, levels = patietOrder)

labs <- c("T(8;21)","Inv(16)/T(16;16)", "FLT3-ITD", "NPM1 Mutation", "CEBPA Mutation", 
          "Monosomy5/Del5q", "Monosomy7", "RAM", "CBFA2T3-GLIS2") %>% set_names(groups)

oncotile <- ggplot(tile.tidy, aes(x=variable, y=CytGroup, fill = value)) + 
  geom_tile(color="white", size=0.5) +
  scale_y_discrete(labels=labs) + 
  labs(title="Co-occuring Cytogenetic Abnormalities", y="", x="") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size=16, color="black"),
        plot.title = element_text(size=18, hjust=0.5))



# tiff(file="~/RNA_seq_Analysis/2017.02.15_CBF-GLIS_DEG/ASH/CBFGLIS_0531_1031_Oncoprint2.tif", height = 6, width = 10, units = "in", res=400)
oncotile
# dev.off()
# getwd()
```



















